### 2025-04-27 paper/paper.tex
Update paper.tex

Our decentralized multi‐agent reinforcement learning (MARL) architecture is built around three key components: the Q‐network itself, a DQN‐based agent wrapper (complete with replay buffer and learning logic), and a SUMO–traci environment interface that supplies state and reward information every control interval. Traci allows us to easily modify our simulation and get state information from within the current timestep, while pytorch allows us to efficiently build and implement our neural networks. This implementation lacks a central coordinator. Each agent acts on its own in coordination with other agents. The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time. We define the neural network in models.py as a simple multilayer perceptron. Given an input dimension equal to the total number of lanes feeding into a signalized intersection and an output dimension equal to the number of possible signal phases, we stack two hidden layers of configurable width (48 units each in our experiments) with ReLU activations, and terminate with a linear layer producing a Q‐value for each phase.

@@

On the environment side (multi\_env.py, subclassing BaseEnv), we employ the python library traci (traffic control interface) to simulate traffic in Starkville, MS. Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and the time frame would not allow us to do this. Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly heavy and prone to congestion. Most of these points were decided by considering the importance of the road, the typical time spent at intersections, and personal experience. This method ensures that we don't expend computational resources at intersections that either see little traffic or do fine managing traffic on their own. Doing so significantly decreases the computational requirements to run the model, and allows us to focus only on critical points. At startup we invoke traci.start with a step length of 0.1s and a control interval of 5s, disabling verbose logs to save overhead. We automatically collect, for each traffic‐light signal (TLS), the set of incoming lanes by inspecting traci.trafficlight.getControlledLinks. Every control step, get\_state() returns a list of queue lengths, i.e.\ the number of halting vehicles per lane, by calling traci.lane.getLastStepHaltingNumber on each lane. We then compute a scalar reward for each TLS as the negative sum of its queue lengths, which incentivizes the agents to minimize congestion. When step(actions) is called, we apply each agent’s chosen phase via traci.trafficlight.setPhase, advance the simulation by the control interval, and then return the new state, reward list, and a done flag (triggered when no vehicles remain).

>>>>

Our decentralized multi‐agent reinforcement learning (MARL) architecture is built around three key components: the Q‐network itself, a Deep Q-Network(DQN)‐based agent wrapper (complete with replay buffer and learning logic), and a SUMO–TraCI environment interface that supplies state and reward information every control interval. Traci allows us to easily modify our simulation and get state information from within the current timestep, while PyTorch allows us to efficiently build and implement our neural networks. This implementation lacks a central coordinator. Each agent acts on its own in coordination with other agents. The agents were designed to represent an intersection with a traffic signal controller, and could accept the state of the lanes leading to the intersection, the state of the signals, and the states of vehicles at each intersection such as stopped time. We define the neural network in models.py as a simple multilayer perceptron. Given an input dimension equal to the total number of lanes feeding into a signalized intersection and an output dimension equal to the number of possible signal phases, we stack two hidden layers of configurable width (48 units each in our experiments) with ReLU activations, and terminate with a linear layer producing a Q‐value for each phase.

@@

On the environment side (multi\_env.py, subclassing BaseEnv), we employ the Python library TraCI (traffic control interface) to simulate traffic in Starkville, MS. Our initial plan was to predict the optimal states of every intersection with a traffic light in Starkville, MS, however, upon implementation we realized our available hardware and the time frame would not allow us to do this. Instead, we opted to only use what we decided were points of interest, or intersections where traffic is particularly heavy and prone to congestion. Most of these points were decided by considering the importance of the road, the typical time spent at intersections, and personal experience. This method ensures that we don't expend computational resources at intersections that either see little traffic or do fine managing traffic on their own. Doing so significantly decreases the computational requirements to run the model, and allows us to focus only on critical points. At startup we invoke traci.start with a step length of 0.1s and a control interval of 5s, disabling verbose logs to save overhead. We automatically collect, for each traffic‐light signal (TLS), the set of incoming lanes by inspecting traci.trafficlight.getControlledLinks. Every control step, get\_state() returns a list of queue lengths, i.e.\ the number of halting vehicles per lane, by calling traci.lane.getLastStepHaltingNumber on each lane. We then compute a scalar reward for each TLS as the negative sum of its queue lengths, which incentivizes the agents to minimize congestion. When step(actions) is called, we apply each agent’s chosen phase via traci.trafficlight.setPhase, advance the simulation by the control interval, and then return the new state, reward list, and a done flag (triggered when no vehicles remain).
